---
title: "Coursera Stats with R Week 1"
output: html_notebook
---

Install course packages
```{r}
install.packages("devtools")
library(devtools)
```
Load other packages (it requires dplyr and ggplot2, which are already installed so I'm skipping them here.)

```{r}
install.packages("shiny")
install_github("StatsWithR/statsr")
```

Now turn on packages:
```{r}
library(dplyr)
library(ggplot2)
library(statsr)
```
Dataset 1. Dr Arbuthnot's baptism records
```{r}
data(arbuthnot)
```
Look at the data
```{r}
arbuthnot
```
View dimensions of the data frame
```{r}
dim(arbuthnot)
```
Check the names of the variables
```{r}
names(arbuthnot)
```
Examine the boys data
```{r}
arbuthnot$boys
```
and the girls...
```{r}
arbuthnot$girls
```
Now create a simple plot of the number of girls baptized per year
```{r}
ggplot(data = arbuthnot, aes(x = year, y = girls))+
  geom_point()
```
Try getting help for ggplot
```{r}
?ggplot
```
Now determine the total number of births per year by adding boys and girls
```{r}
arbuthnot$boys+arbuthnot$girls
```
Now use dplyr to add a new variable to the dataframe
```{r}
arbuthnot<-arbuthnot %>% mutate(total = boys + girls)
```
Now plot total per year:
```{r}
ggplot(arbuthnot, aes(year, total))+
  geom_line()+
  geom_point()
```
Now see if you can plot the proportion of boys born over time
```{r}
ggplot(arbuthnot, aes(year, (boys/total)*100))+
  geom_point()+
  geom_line()
```
Now let's use R to make comparisons

Do boys outnumber girls each year?
```{r}
arbuthnot<-arbuthnot %>% 
  mutate(more_boys = boys > girls)
```
Dataset 2. Present birth records
Load a dataset
```{r}
data(present)
```
What years are included in the dataset? Use the range() function.

```{r}
range(present$year)
```
Now calculate the total number of births for each year and store in new variable called `total`.  Then calculate the proportion of boys (call it prop_boys).  Then plot to determine if the proportion of boys born in the US has decreased over time.

```{r}
present <- present %>% 
  mutate(total = boys + girls,
         prop_boys = (boys/total)*100)

ggplot(present, aes(year, prop_boys))+
  geom_line()+
  geom_point()
```
Now see for which years there were more boys than girls
```{r}
present<- present %>% 
  mutate(more_boys = boys > girls)
```
Now calculate the boy:girl ratio and store it in a variable called prop_boy_girl and then plot over time.

```{r}
present <- present %>% 
  mutate(prop_boy_girl = boys/girls)

ggplot(present, aes(year, prop_boy_girl))+
  geom_line()+
  geom_point()
```
Now arrange the data in descending order in order to see the year with the highest total births using the `desc()` function:
```{r}
desc(present$total)
```


